---
output: github_document
---


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# ccharter

Control Charts made easy. ccharter focus on calculating control systems in a time series and plotting them as an overlay to the series.

# Installation

* You can install from github with:

    ```R
    if ( !("devtools" %in% installed.packages()) ) {
       install.packages("devtools")
    }
    devtools::install_github("jbaxx/ccharter")
    ```
# Using ccharter

ccharter comes with two functions: 

*  `ccpoints` to calculate Control Systems points from a time series data frame
*  `cc2plot` to plot the Control Systems points (Control Chart)

Loading the package

```{r}
library(ccharter)
```

Calculating the Control Systems points

* An example usage of `ccpoints`:

```{r}
library(ccharter)
# Create a random time series
set.seed(154)
time.series <- data.frame(t.dates = seq.Date(as.Date("2014-02-01"), as.Date("2016-08-01"), "month"),
               t.values = c(
               seq(0.1, 0.8, by = 0.1) * runif(8) + 3,
               seq(0.1, 0.7, by = 0.1) * runif(7) + 4,
               seq(0.1, 0.7, by = 0.1) * runif(7) + 5,
               seq(0.1, 0.4, by = 0.1) * runif(4) + 4,
               seq(0.1, 0.5, by = 0.1) * runif(5) + 4)
               )

# Execute function
control.chart.data <- ccpoints(time.series, "t.dates", "t.values")
print(control.chart.data)

# To extract only the data frame
control.chart.data[["data"]]
```


Plotting the Control Systems points

* An example usage of `cc2plot`:

```{r example, fig.width=9,fig.height=4}
library(ccharter)
# Create a random time series
set.seed(154)
time.series <- data.frame(t.dates = seq.Date(as.Date("2014-02-01"), as.Date("2016-08-01"), "month"),
               t.values = c(
               seq(0.1, 0.8, by = 0.1) * runif(8) + 3,
               seq(0.1, 0.7, by = 0.1) * runif(7) + 4,
               seq(0.1, 0.7, by = 0.1) * runif(7) + 5,
               seq(0.1, 0.4, by = 0.1) * runif(4) + 4,
               seq(0.1, 0.5, by = 0.1) * runif(5) + 4)
               )

# Execute function
control.chart.data <- ccpoints(time.series, "t.dates", "t.values")

# Create chart
cc2plot(control.chart.data)
```


For detailed functions documentation you can access the help files in R with `?ccpoints` and `?cc2plot`
